RDF Validation Requirements - Evaluation and Logical Underpinning

### 2015-01-05 paper/RDF validation requirements/RDF validation requirements.tex
finalize RDF validation requirements paper

\newenvironment{gcotable-old}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}\linebreak
  \begin{tabular}{l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context concept} & \textbf{p. list} & \textbf{concepts} & \textbf{c. element} & \textbf{c. value} \\
  \hline

@@

\newenvironment{gcotable}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
  \hline

@@

\newenvironment{DL}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
	\begin{center}
  \begin{tabular}{r l}

@@

\newenvironment{evaluation}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{constraint} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

@@

\newenvironment{evaluation-overall}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{constraint} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

@@

\newenvironment{evaluation-generic}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|c|c|c}
  \hline
  \textbf{constraint} & \textbf{property c.} & \textbf{simple c.} & \textbf{DL} \\
  \hline

@@

\newenvironment{evaluation-generic-overview}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|c|c}
  \hline
  \textbf{Constraint Classes} & \textbf{\#} & \textbf{\%} \\
  \hline

@@

\ms{Subsumption} corresponds to the requirement \ms{R-100-SUBSUMPTION}.

\begin{itemize}
	\item DL terminology: concept inclusion
\end{itemize}

\begin{itemize}
	\item all mothers are parents
  \item the concept Mother is subsumed by the concept Parent
\end{itemize}

@@

Jedis feel the force.

@@

\begin{ex}
# OWL2
# ----
Jedi rdfs:subClassOf FeelingForce .
\end{ex}

\begin{ex}
# ReSh
# ----
the extension ext:extendsShape may be used
\end{ex}

\begin{ex}
# ShEx
# ----
FeelingForce {
    feelingForce (true) }
Jedi {
    & FeelingForce ,
    attitute ('good') }
\end{ex}

\begin{ex}
# data
# ----
Yoda 
    feelingForce true ;
    attitute 'good' .
\end{ex}

@@

@@

\begin{itemize}
	\item R-3-EQUIVALENT-CLASSES
\end{itemize}

Concept equivalence asserts that two concepts have the same instances \cite{Kroetzsch2012}.
While synonyms are an obvious example of equivalent concepts, in practice one more
often uses concept equivalence to give a name to complex expression  \cite{Kroetzsch2012}.
Concept equivalence is indeed subsumption from left and right ($A \sqsubseteq B$ and $B \sqsubseteq A$ implies $A \equiv B$ ).

@@

\section{Sub Object Properties}

\begin{itemize}

	\item R-54-SUB-OBJECT-PROPERTIES
\end{itemize}

\begin{itemize}
	\item DL terminology: role inclusion 
\end{itemize}

@@

\section{Sub Data Properties}

\begin{itemize}

	\item R-54-SUB-DATA-PROPERTIES
\end{itemize}

\begin{itemize}
	\item DL terminology: concrete domains functional roles  
\end{itemize}

\subsection{Complex Example}
\begin{DL}
 $Person \sqcap  \forall hasAge. \leq_{10} \sqsubseteq Person \sqcap \forall hasEvenAge. \leq_{10}$ 
\end{DL}

\section{Object Property Paths / Object Property Chains}

\begin{itemize}
	\item R-55-OBJECT-PROPERTY-PATHS
\end{itemize}

\begin{itemize}
  \item DL terminology: complex role inclusion axiom / role composition
	\item role composition can only appear on the left-hand side of complex role inclusions \cite{Kroetzsch2012}
\end{itemize}

@@

\begin{ex}
# OWL2
# ----
:uncleOf owl:propertyChainAxiom ( :brotherOf :parentOf ) . 
\end{ex}

@@

\begin{itemize}
	\item R-9-DISJOINT-PROPERTIES
\end{itemize}

\begin{DL}
Disjoint(parentOf, childOf)\\
Alternative:\\
$parentOf \sqsubseteq \neg childOf$
\end{DL}

@@

syntactic sugar (preferred):

@@

\begin{itemize}
  \item R-15-CONJUNCTION-OF-CLASS-EXPRESSIONS
	\item R-16-CONJUNCTION-OF-DATA-RANGES
\end{itemize}

\begin{itemize}
	\item composition / conjunction
\end{itemize}

DLs allow new concepts and roles to be
built using a variety of different constructors. We distinguish concept and role 
constructors depending on whether concept or role expressions are constructed. In the case of
concepts, one can further separate basic Boolean constructors, role restrictions and nominals/enumerations \cite{Kroetzsch2012}.

Boolean concept constructors provide basic boolean operations that are closely related to
the familiar operations of intersection, union and complement of sets, or to conjunction,
disjunction and negation of logical expressions \cite{Kroetzsch2012}.

@@

concept inclusions allow us to state that all mothers are female and that
all mothers are parents, but what we really mean is that mothers are exactly the female
parents. DLs support such statements by allowing us to form complex concepts such as
the intersection (also called conjunction)
which denotes the set of individuals that are both female and parents. A complex concept
can be used in axioms in exactly the same way as an atomic concept, e.g., in the
equivalence Mother $\equiv$ Female $\sqcap$ Parent .

@@

Complex concept of all individuals which are of the concept Female and of the concept Parent.

@@

\begin{itemize}
	\item R-17-DISJUNCTION-OF-CLASS-EXPRESSIONS
	\item R-18-DISJUNCTION-OF-DATA-RANGES
\end{itemize}

\begin{itemize}
	\item synonyms: union / inclusive OR
	\item disjunction of classes, data ranges
\end{itemize}

@@

\begin{itemize}
  \item R-19-NEGATION-OF-CLASS-EXPRESSIONS
	\item R-20-NEGATION-OF-DATA-RANGES
\end{itemize}

\begin{itemize}
	\item synonym: complement
\end{itemize}

@@

set of all individuals that are not married.

@@

@@

requirement \ms{R-7-DISJOINT-CLASSES}

\begin{DL}
Male $\sqcap$ Female $\sqsubseteq$ $\perp$\\
Alternative:\\
$Male \sqsubseteq \neg Female$
\end{DL}

@@

\begin{DL}
Hologram $\sqcap$ Human $\sqsubseteq$ $\perp$\\
Alternative:\\
$Hologram \sqsubseteq \neg Human$
\end{DL}

@@

\ms{Existential Quantification on Properties} conforms to the requirement \ms{R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}
In DL terminology the \ms{existential quantification} is alco called \ms{existential restriction}.
An existential class expression ObjectSomeValuesFrom( OPE CE ) consists of an object property expression OPE and a class expression CE, and it contains all those individuals that are connected by OPE to an individual that is an instance of CE. Such a class expression can be seen as a syntactic shortcut for the class expression ObjectMinCardinality( 1 OPE CE ).
An existential class expression DataSomeValuesFrom( DPE1 ... DPEn DR ) consists of n data property expressions DPEi, 1 $\leq$ i $\leq$ n, and a data range DR whose arity MUST be n. Such a class expression contains all those individuals that are connected by DPEi to literals lti, 1 $\leq$ i $\leq$ n, such that the tuple ( lt1 , ..., ltn ) is in DR. A class expression of the form DataSomeValuesFrom( DPE DR ) can be seen as a syntactic shortcut for the class expression DataMinCardinality( 1 DPE DR ).

@@

complex concept that describes the set
of individuals that are parents of at least one individual (instance of $\top$).

@@

the complex concept describes those individuals that are parents of at least one
female individual, i.e., those that have a daughter.

@@

\ms{Universal Quantification on Properties} corresponds to the requirement \ms{R-91-UNIVERSAL-QUANTIFICATION-ON-PROPERTIES}, which is called \ms{value restriction} in DL terminology.

@@

\begin{itemize}
	\item set of individuals all of whose children are female
  \item also includes those that have no children at all
\end{itemize}

@@

\begin{itemize}
	\item R-25-OBJECT-PROPERTY-DOMAIN
	\item R-26-DATA-PROPERTY-DOMAIN
\end{itemize}

\begin{itemize}
	\item restrict domain of object and data properties
	\item DL terminology: domain restrictions on roles
	\item The language needs to include an easy-to-use vocabulary to declare that a given property is associated with a class, e.g. to populate input forms with appropriate widgets but also constraint checking. In OO terms this is the declaration of a member, field, attribute or association. 
\end{itemize}

$\exists R. \top \sqsubseteq C$ is the object property restriction where $R$ is the object property (role) whose domain is restricted to concept $C$.

@@

\begin{itemize}
	\item restrict the domain of sonOf to male individual
\end{itemize}

@@

\begin{itemize}
	\item R-28-OBJECT-PROPERTY-RANGE
	\item R-35-DATA-PROPERTY-RANGE
\end{itemize}

\begin{itemize}
	\item restrict range of object and data properties
	\item DL terminology: range restrictions on roles
\end{itemize}

$\top \sqsubseteq \forall R . C$ is the range restriction to object property $R$ (restricted by concept $C$).

\begin{DL}
$\top$ $\sqsubseteq$ $\forall$ sonOf . Parent \\
Equivalent to:\\
$\exists sonOf\sp{\overline{\ }}.\top \sqsubseteq Person$
\end{DL}

\begin{itemize}
	\item restrict the range of sonOf to parents
\end{itemize}

\begin{ex}
# OWL 2
# -----
sonOf rdfs:range Parent . 
\end{ex}

\begin{ex}
# DSP
# ---
:hasDogRange
        a dsp:DescriptionTemplate ; 
        dsp:resourceClass owl:Thing ; 
        dsp:statementTemplate [
            a dsp:NonLiteralStatementTemplate ;
            dsp:property sonOf ; 
            dsp:nonLiteralConstraint [ 
                a dsp:NonLiteralConstraint ;
                dsp:valueClass Parent ] ] .
\end{ex}

@@

\begin{itemize}
	\item R-29-CLASS-SPECIFIC-RANGE-OF-RDF-OBJECTS
	\item R-36-CLASS-SPECIFIC-RANGE-OF-RDF-LITERALS
\end{itemize}

\begin{itemize}
  \item restrict range of object and data properties for individuals within specific context
	\item context: e.g. class
	\item The values of each member property of a class may be limited by their value type, such as xsd:string or ex:Person. 
\end{itemize}

\begin{itemize}
	\item property 'sonOf'
	\item context: class 'Man'
	\item range: class 'Parent'
	\item $\rightarrow$ individuals of the class 'Man' can only have 'sonOf' relationships to indivuduals of the class 'Parent'
\end{itemize}

@@

@@

\begin{itemize}
	\item R-81-MINIMUM-UNQUALIFIED-CARDINALITY-ON-PROPERTIES
	\item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

$\leq n R. \top$ is minimum unqualified cardinality restriction where $n \in \mathbb{N}$ (written $\leq  n R$ in short).

\begin{itemize}
	\item A minimum cardinality expression ObjectMinCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at least n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A minimum cardinality expression DataMinCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at least n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item unqualified: CE respective DR is NOT stated. 
\end{itemize}

@@

\begin{ex}
# Individuals matching the ’:SuperJediMaster’ data shape
# ------------------------------------------------------
:Yoda 
\end{ex}

@@

>>>>

\newenvironment{gcotable-old}{
  \scriptsize
  \sffamily
  \vspace{0cm}\linebreak
  \begin{tabular}{l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context concept} & \textbf{p. list} & \textbf{concepts} & \textbf{c. element} & \textbf{c. value} \\
  \hline

@@

\newenvironment{gcotable}{
  \scriptsize
  \sffamily
  \vspace{0cm}
	\begin{center}
	\textbf{\vspace{0.4cm}Mapping to RDF-CO:} \\
  \begin{tabular}{l|l|l|l|l|l|l}
	\hline
  \textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
  \hline

@@

\newenvironment{DL}{
  %\scriptsize
  %\sffamily
  \vspace{0cm}
	\begin{center}
  \begin{tabular}{r l}

@@

\newenvironment{evaluation}{
  \scriptsize
  \sffamily
  \vspace{0cm}
	\begin{center}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{constraint} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

@@

\newenvironment{evaluation-overall}{
  \scriptsize
  \sffamily
  \vspace{0cm}
	\begin{center}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{constraint} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

@@

\newenvironment{evaluation-generic}{
  \scriptsize
  \sffamily
  \vspace{0cm}
	\begin{center}
  \begin{tabular}{l|c|c|c}
  \hline
  \textbf{constraint} & \textbf{property c.} & \textbf{simple c.} & \textbf{DL} \\
  \hline

@@

\newenvironment{evaluation-generic-overview}{
  \scriptsize
  \sffamily
  \vspace{0cm}
	\begin{center}
  \begin{tabular}{l|c|c}
  \hline
  \textbf{Constraint Classes} & \textbf{\#} & \textbf{\%} \\
  \hline

@@

\ms{Subsumption} (DL terminology: \ms{concept inclusion}) corresponds to the requirement \ms{R-100-SUBSUMPTION}.
A subclass axiom SubClassOf( CE1 CE2 ) states that the class expression CE1 is a subclass of the class expression CE2. 
Roughly speaking, this states that CE1 is more specific than CE2.

All mothers are parents.
The concept \ms{Mother} is subsumed by the concept \ms{Parent}:

@@

Jedis feel the force:

@@

Expressed by multiple constraint languages:

\begin{ex}
# OWL2:
Jedi rdfs:subClassOf FeelingForce .
\end{ex}

\begin{ex}
# ReSh:
the extension ext:extendsShape may be used
\end{ex}

\begin{ex}
# ShEx:
FeelingForce {
    feelingForce (true) }
Jedi {
    & FeelingForce ,
    attitute ('good') }
\end{ex}

Data matching the shapes \ms{FeelingForce} and \ms{Jedi}:

\begin{ex}
Yoda 
    feelingForce true ;
    attitute 'good' .
\end{ex}

@@

If an individual is rich, then this individual is not poor:

@@

\ms{Class Equivalence} corresponds to the requirement \ms{R-3-EQUIVALENT-CLASSES}.
Concept equivalence asserts that two concepts have the same instances \cite{Kroetzsch2012}.
While synonyms are an obvious example of equivalent concepts, in practice one more
often uses concept equivalence to give a name to complex expressions \cite{Kroetzsch2012}.
Concept equivalence is indeed subsumption from left and right ($A \sqsubseteq B$ and $B \sqsubseteq A$ implies $A \equiv B$ ).

@@

\section{Sub Properties}

\ms{Sub Properties} (DL terminology: \ms{role inclusion}) correspond to the requirements \ms{R-54-SUB-OBJECT-PROPERTIES} and \ms{R-54-SUB-DATA-PROPERTIES}.
Subproperty axioms are analogous to subclass axioms.
These axioms state that the property expression PE1 is a subproperty of the property expression PE2 — that is, if an individual x is connected by PE1 to an individual or a literal y, then x is also connected by PE2 to y.

@@

%\begin{itemize}
	%\item DL terminology: concrete domains functional roles  
%\end{itemize}

\begin{DL}
 $Person \sqcap  \forall hasAge. \leq_{10} \sqsubseteq Person \sqcap \forall hasEvenAge. \leq_{10}$ 
\end{DL}

\section{Object Property Paths}

\ms{Object Property Paths} (or Object Property Chains and in DL terminology complex role inclusion axiom or role composition)
corresponds to the requirement \ms{R-55-OBJECT-PROPERTY-PATHS}.
The more complex form of sub properties. This axiom states that, if an individual x is connected by a sequence of object property expressions OPE1, ..., OPEn with an individual y, then x is also connected with y by the object property expression OPE.  
Role composition can only appear on the left-hand side of complex role inclusions \cite{Kroetzsch2012}.

@@

\begin{ex}
# OWL2:
uncleOf owl:propertyChainAxiom ( brotherOf parentOf ) . 
\end{ex}

@@

\ms{Disjoint Properties} corresponds to the requirement \ms{R-9-DISJOINT-PROPERTIES}.
A disjoint properties axiom states that all of the properties are pairwise disjoint; 
that is, no individual x can be connected to an individual y by these properties.

The object properties \ms{parentOf} and \ms{childOf} are disjoint:

\begin{DL}
Disjoint(parentOf, childOf)\\
\end{DL}

or alternatively:

\begin{DL}
$parentOf \sqsubseteq \neg childOf$
\end{DL}

@@

@@

\ms{Intersection} (composition, conjunction) corresponds to the requirements \ms{R-15-CONJUNCTION-OF-CLASS-EXPRESSIONS} and \ms{R-16-CONJUNCTION-OF-DATA-RANGES}.
DLs allow new concepts and roles to be
built using a variety of different constructors. We distinguish concept and role 
constructors depending on whether concept or role expressions are constructed. In the case of
concepts, one can further separate basic Boolean constructors, role restrictions and nominals/enumerations \cite{Kroetzsch2012}.
Boolean concept constructors provide basic boolean operations that are closely related to
the familiar operations of intersection, union and complement of sets, or to conjunction,
disjunction and negation of logical expressions \cite{Kroetzsch2012}.

@@

Concept inclusions allow us to state that all mothers are female and that
all mothers are parents, but what we really mean is that mothers are exactly the female
parents. DLs support such statements by allowing us to form complex concepts such as
the intersection (also called conjunction)
which denotes the set of individuals that are both female and parents. A complex concept
can be used in axioms in exactly the same way as an atomic concept, e.g., in the
equivalence Mother $\equiv$ Female $\sqcap$ Parent .

@@

Complex concept of all individuals which are of the concept \ms{Female} and of the concept \ms{Parent}.

@@

\ms{Disjunction} of classes or data ranges corresponds to the requirements \\
\ms{R-17-DISJUNCTION-OF-CLASS-EXPRESSIONS} and \ms{R-18-DISJUNCTION-OF-DATA-RANGES}.
Synonyms are union and inclusive or.

@@

\ms{Negation} (complement) corresponds to the requirements \ms{R-19-NEGATION-OF-CLASS-EXPRESSIONS} and \ms{R-20-NEGATION-OF-DATA-RANGES}.

@@

Set of all individuals that are not married.

@@

All female individuals that are not married.

@@

\ms{Disjoint Classes} corresponds to the requirement \ms{R-7-DISJOINT-CLASSES}.
A disjoint classes axiom states that all of the classes are pairwise disjoint; 
that is, no individual can be at the same time an instance of these classes.

Individuals cannot be male and female at the same time:

\begin{DL}
Male $\sqcap$ Female $\sqsubseteq$ $\perp$\\
\end{DL}

or alternatively:

\begin{DL}
$Male \sqsubseteq \neg Female$
\end{DL}

@@

One can either be a hologram or a human, but not both:

\begin{DL}
Hologram $\sqcap$ Human $\sqsubseteq$ $\perp$\\
\end{DL}

or alternatively:

\begin{DL}
$Hologram \sqsubseteq \neg Human$
\end{DL}

@@

\ms{Existential Quantification on Properties} conforms to the requirement \\
\ms{R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}.
In DL terminology the \ms{existential quantification} is alco called \ms{existential restriction}.
An existential class expression consists of a property expression and a class expression or a data range, and it contains all those individuals that are connected by the property expression to an individual that is an instance of the class expression or to literals that are in the data range.

@@

Complex concept that describes the set
of individuals that are parents of at least one individual (instance of $\top$).

@@

The complex concept describes those individuals that are parents of at least one
female individual, i.e., those that have a daughter.

@@

\ms{Universal Quantification on Properties} corresponds to the requirement \ms{R-91-UNIVERSAL-QUANTIFICATION-ON-PROPERTIES}, 
which is also called \ms{value} \ms{restriction} in DL terminology.

@@

The set of individuals all of whose children are female also includes those that have no children at all.

@@

\ms{Property Domain} corresponds to the requirements \ms{R-25-OBJECT-PROPERTY-DOMAIN} and \ms{R-26-DATA-PROPERTY-DOMAIN}.
The constraint restricts the domain of object and data properties.
In DL terminology this constraint is also called domain restrictions on roles.
The purpose is to declare that a given property is associated with a class, e.g. to populate input forms with appropriate widgets but also constraint checking. In OO terms this is the declaration of a member, field, attribute or association. 
$\exists R. \top \sqsubseteq C$ is the object property restriction where $R$ is the object property (role) whose domain is restricted to concept $C$.

@@

Restricts the domain of \ms{sonOf} to male individual.

@@

\ms{Property Range} corresponds to the requirements \ms{R-28-OBJECT-PROPERTY-RANGE} and \ms{R-35-DATA-PROPERTY-RANGE}.
This constraint restricts the range of object and data properties
In DL terminology it is also called range restrictions on roles.
$\top \sqsubseteq \forall R . C$ is the range restriction to the object property $R$ (restricted by the concept $C$).

\begin{DL}
$\top$ $\sqsubseteq$ $\forall$ sonOf . Parent \\
\end{DL}

equivalent to:

\begin{DL}
$\exists sonOf\sp{\overline{\ }}.\top \sqsubseteq Person$
\end{DL}

Restricts the range of \ms{sonOf} to parents.

\begin{ex}
# OWL 2:
sonOf rdfs:range Parent . 
\end{ex}

\begin{ex}
# DSP:
:hasDogRange
        a dsp:DescriptionTemplate ; 
        dsp:resourceClass owl:Thing ; 
        dsp:statementTemplate [
            a dsp:NonLiteralStatementTemplate ;
            dsp:property sonOf ; 
            dsp:nonLiteralConstraint [ 
                a dsp:NonLiteralConstraint ;
                dsp:valueClass Parent ] ] .
\end{ex}

@@

\ms{Class-Specific Property Range} corresponds to the requirements
\ms{R-29-CLASS-} \ms{SPECIFIC-RANGE-OF-RDF-OBJECTS} and \ms{R-36-CLASS-SPECIFIC-RANGE-OF-RDF-} \ms{LITERALS}.
The constraint restricts the range of object and data properties for individuals within a specific context (e.g. class, application profile).
The values of each member property of a class may be limited by their value type, such as xsd:string or Person.

@@

Individuals of the class \ms{Man} (context class) can only have \ms{sonOf} relationships to individuals of the class \ms{Parent} (range class).

@@

\ms{Minimum Unqualified Cardinality Restrictions on Properties}
corresponds to the requirements
\ms{R-81-MINIMUM-UNQUALIFIED-CARDINALITY-ON-PROPERTIES} and
\ms{R-211-CARDINALITY-CONSTRAINTS}.
$\leq n R. \top$ is the minimum unqualified cardinality restriction where $n \in \mathbb{N}$ (written $\leq  n R$ in short).
A minimum cardinality restrictions contains all those individuals that are connected by a property to at least n different individuals/literals 
that are instances of a particular class or data range. If the class is missing, it is taken to be owl:Thing. 
If the data range is missing, it is taken to be rdfs:Literal.

\begin{ex}
# ShEx:
:Jedi {
    :attitude ('good') }
:JediStudent {
    & :Jedi ,
    :studentOf {}{1} }
:JediMaster {
    & :Jedi ,
    :mentorOf {}{1,2} }
:SuperJediMaster {
    & :Jedi ,
    :mentorOf {}{3,} }
\end{ex}

@@

\begin{ex}
# data:
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:SuperJediMaster’ data shape:
:Yoda 
\end{ex}

@@

Captains command at least one vessel.
